A boundary integral equation approach to computing eigenvalues of the Stokes operator

### 2019-03-26 doc/draft-01-stokes/01intro.tex
add some small edits

The eigenvalues (and eigenfunctions)
of the Stokes operator have applications in the
stability analysis of stationary solutions of the
Navier--Stokes equations \cite{osborn1976approximation}
and in the study of decaying two dimensional turbulence
\cite{schneider2008final}.
%
The eigenvalues and eigenfunctions of the Stokes
operator are the subject of intense analytical
investigation
\cite{taylor1933buckling,szego1950membranes,
  polya1951isoperimetric,bramble1963pointwise,
  ashbaugh1996fundamental,leriche2004stokes,
  kelliher2009eigenvalues,antunes2011buckling},
particularly how they relate to the eigenvalues and
eigenfunctions of the drum.

Historically, the Stokes eigenvalue problem serves as a
common model problem for numerical eigenvalue analysis
with a fourth order operator (here, the bi-Laplacian).
%
Further, numerical simulation has long played an
important role in the analyses described above --- both for
estimating the eigenvalues on domains of practical
interest and for exploring open conjectures.

Borrowing the language of~\cite{zhao2015robust},
which concerns the Laplace or ``drum'' eigenvalue
problem,
the numerical treatment of the Stokes eigenvalue
problem can be divided into two basic approaches.
%
The first class of methods
directly discretize the
Stokes operator, typically with a
finite element basis, and the eigenvalues are found
as the eigenvalues of the discrete system.
%
The second class of methods reformulate the 
oscillatory Stokes equations as a boundary integral
equations which are discretized.
The eigenvalues are then found
by a nonlinear search for the values of $k$ where the BIE 
is not invertible.
%

@@

While integral equation based methods for the
related ``buckling'' eigenvalue problem
(which is equivalent  on simply connected domains
\cite{kelliher2009eigenvalues})
have been considered previously,
these typically relied on first-kind integral
equation formulations of the underlying PDE,
i.e. formulations in which the BIE operator is
compact \cite{kitahara2014boundary,antunes2011buckling}.
%
This is unsatisfying from a theoretical
perspective, because compact operators always
have nontrivial nullspaces
(though, by design, the representations in
\cite{kitahara2014boundary,antunes2011buckling}
are not injective precisely when
$k^2$ is an eigenvalue).
%
This obscures the relation between
the non-invertibility of the
discrete system and the eigenvalues;
in particular, common measures of the
``non-invertibility'' of a matrix, like the
smallest singular value or the determinant,
converge rapidly to zero for all values
of $k^2$ as the boundary is refined.
The measure of whether $k^2$ is an
approximate eigenvalue is then {\em relative to the
  current grid} for first kind formulations.

>>>>

The eigenvalues (and eigenfunctions)
of the Stokes operator have applications in the
stability analysis of stationary solutions of the
Navier--Stokes equations \cite{osborn1976approximation}
and in the study of decaying two dimensional turbulence
\cite{schneider2008final}.
%
The eigenvalues and eigenfunctions of the Stokes
operator are the subject of intense analytical
investigation
\cite{taylor1933buckling,szego1950membranes,
  polya1951isoperimetric,bramble1963pointwise,
  ashbaugh1996fundamental,leriche2004stokes,
  kelliher2009eigenvalues,antunes2011buckling},
especially as they relate to the eigenvalues and
eigenfunctions of the Laplacian.

Historically, the Stokes eigenvalue problem serves as a
common model problem for numerical eigenvalue analysis
with a fourth order operator (here, the bi-Laplacian).
%
Further, numerical simulation has long played an
important role in the analyses cited above --- both for
estimating the eigenvalues on domains of practical
interest and in forming new conjectures.

Borrowing the language of~\cite{zhao2015robust},
which concerns the eigenvalues of the Laplacian
(also known as the membrane or ``drum'' problem),
the numerical treatment of the Stokes eigenvalue
problem can be divided into two basic approaches.
%
The first class of methods
directly discretize the
Stokes operator, typically with a
finite element basis, and the eigenvalues are found
as the eigenvalues of the discrete system.
%
The second class of methods reformulate the 
oscillatory Stokes equations as a boundary integral
equations which are discretized.
The eigenvalues are then found
by a nonlinear search for the values of $k$ where the BIE 
is not invertible.
%

@@

While integral equation based methods for the
related ``buckling'' eigenvalue problem
(which is equivalent  on simply connected domains
\cite{kelliher2009eigenvalues})
have been considered previously,
these typically relied on first-kind integral
equation formulations of the underlying PDE,
i.e. formulations in which the BIE operator is
compact \cite{kitahara2014boundary,antunes2011buckling}.
%
This is unsatisfying from a numerical
perspective, because the spectrum of a
compact operator either contains zero
or has zero as a limit point
(by design, the representations in
\cite{kitahara2014boundary,antunes2011buckling}
are not injective precisely when
$k^2$ is an eigenvalue).
%
This obscures the relation between
the non-invertibility of discrete approximations
of the operator and the eigenvalues;
in particular, common measures of the
``non-invertibility'' of a matrix, like the
smallest singular value or the determinant,
converge rapidly to zero for all values
of $k^2$ as the boundary is refined.
The measure of whether $k^2$ is an
approximate eigenvalue is then {\em relative to the
  current grid} for first kind formulations.

### 2019-03-26 doc/draft-01-stokes/05conclusion.tex
add some small edits

\begin{equation*}
  \psi = \partial_\nu \psi = 0 \; \quad \textrm{on} \quad \Gamma.
\end{equation*}
On a multiply connected domain, it is possible for
an eigenfield to have no corresponding stream
function which is still clamped, so that the
buckling eigenvalues are a subset of the Stokes
eigenvalues.
By adapting the approach of \cite{rachh2017integral},
a suitable layer potential representation of the
buckling problem can be derived based on the
oscillatory Stokes layer potentials.
%
This is the subject of a follow-up paper which is in
preparation.

>>>>

\begin{equation*}
  \psi = \partial_\nu \psi = 0 \; \quad \textrm{on} \quad \Gamma.
\end{equation*}
On a multiply connected domain, there are
Stokes eigenfunctions which do not have a corresponding
clamped stream function, so that the
buckling eigenvalues are a subset of the Stokes
eigenvalues.
By adapting the approach of \cite{rachh2017integral},
a suitable layer potential representation of the
buckling problem can be derived based on the
oscillatory Stokes layer potentials.
%
This is the subject of a follow-up paper which is in
preparation.
